Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-05-29 neurips-20/cvx-flats.tex
geez I wasn't even done...

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
	For all $r\in\gamma(p)$, if there is no flat $F$ of codimension at most $d$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.\jessiet{Do we still want to state it as the contrapositive?  That's where the bound comes from, but we use the original direction for all of the intuitive explanation.}
\end{theorem}
\begin{proof}[Proof]
	We prove the contrapositive: if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $d$-codimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

@@

In the proof of Theorem~\ref{thm:consistent-implies-indir-elic} below, we modify the argument of~\citet[Corollary 7]{frongillo2018elicitation} to bound the elicitation complexity of the Bayes risk of the loss.
\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some $\Gamma:\simplex\to\reals^d$.
  Let $p\in\relint\simplex$ and let $\Gamma_r$ be some level set of $\Gamma$ such that (i) $p\in\Gamma_r$, (ii) $\codim(\affhull(\Gamma_r))=d+1$, and (iii) if $\Gamma_r$ is not a singleton, $\lbar$ is nonconstant on $\Gamma_r$.
  Then $\eliccvx(\lbar) \geq \min(d+1,n-1)$.
\end{theorem}
\begin{proof}
  Suppose that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^k$ and link $\psi : \reals^k \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  % The condition $\iden(\Gamma)=d$ implies the existence of some level set $\Gamma_r$ such that $\codim(\affhull(\Gamma_r)) \geq d$; otherwise, taking $F_r = \affhull(\Gamma_r)$ for all reports $r\in\reals^d$, we would have $\codim(F_r) \leq d-1$ and $\Gamma_r = F_r \cap \simplex$, implying $\iden(\Gamma)\leq d-1$.
  The proof of \citet[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that every level set of $\hat\Gamma$ is contained in a level set of $\Gamma$; for completeness we give the argument here.
  Suppose for a contradiction that we have $p,p'$ with $\hat\Gamma(p)=\hat\Gamma(p')$ but $\Gamma(p) \neq \Gamma(p')$.
  As $\lbar = \psi \circ \hat\Gamma$, we also have $\lbar(p) = \lbar(p')$.
  Letting $p'' = \tfrac 1 2 p +  \tfrac 1 2 p'$, Lemma~\ref{lem:elic-complex-bayes-concave} would then give us $\lbar(p'') >  \tfrac 1 2 \lbar(p) +  \tfrac 1 2 \lbar(p') = \lbar(p)$.
  By \citet{osband1985providing}, the level sets $\hat\Gamma_{\hat r}$ are convex, giving $\hat\Gamma(p'') = \hat\Gamma(p)$, which would imply $\lbar(p'')=\lbar(p)$, contradicting $\lbar = \psi \circ \hat\Gamma$.
  We conclude $\hat\Gamma$ must refine $\Gamma$, and in particular, we must have $p \in \hat\Gamma_{\hat r} \subseteq \Gamma_r$ for some $\hat r\in\hat\Gamma(\simplex)$.
  % The proof of \cite[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that for all $\hat r \in \hat\Gamma(\simplex)$ we have $\hat\Gamma_{\hat r} \subseteq \Gamma_r$ for some $r\in\reals^k$.

  Let $F = \affhull(\Gamma_r)$ and recall $\codim(F)=d+1$.
  By Theorem~\ref{thm:cvx-flats}, as $\hat\Gamma$ is $k$-convex elicitable, $\hat\Gamma_{\hat r}$ is the union of flats of codimension at most $k+1$ interesected with $\simplex$; let $\hat F$ be such a flat containing $p$, and $S = \hat F \cap \simplex$, so that $\codim(\hat F) \leq k+1$ and $p \in \hat F\cap\relint\simplex$ and $S \subseteq \hat\Gamma_{\hat r} \subseteq \Gamma_r$.
  As $p$ shows $\relint\simplex\cap \hat F \neq \emptyset$, Lemma~\ref{lem:affhull-relint} gives $\affhull(S) = \affhull(\hat F\cap \simplex) = \hat F$.
  Now as $S \subseteq \Gamma_r$, we also have $\hat F = \affhull(S) \subseteq \affhull(\Gamma_r) = F$,
  implying $\codim(\hat F) \geq \codim(F)$.

  If $\Gamma_r = \{p\}$ is a singleton, then $F = \affhull(\Gamma_r) = \{p\}$, and in particular $n = \codim(F) \leq \codim(\hat F)$, so we must have $k=d=n-1$.
  If $\Gamma_r$ is not a singleton, suppose for a contradiction that $k\leq d$.
  Then $\codim(F) = d+1 \geq k+1 \geq \codim(\hat F)$, so we must have $\codim(\hat F)=d+1$ and $\hat F = F$.
  We now have $S = \hat F \cap \simplex = F \cap \simplex = \Gamma_r$, and as $S \subseteq \hat\Gamma_{\hat r} \subseteq \Gamma_r$, we conclude $\hat\Gamma_{\hat r} = \Gamma_r$.
  By assumption, $\lbar$ is non-constant on $\Gamma_r$, so we have distributions $p,p' \in\hat\Gamma_{\hat r}$ with $\lbar(p)\neq\lbar(p')$, which contradicts $\lbar = \psi \circ \hat\Gamma$.
\end{proof}

>>>>

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
	For all $r\in\gamma(p)$, if there is no flat $F$ of codimension at most $d$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.\jessiet{Do we still want to state it as the contrapositive?  That's where the bound comes from, but we use the original direction for all of the intuitive explanation.}
  \raft{I kind of see your point.  In fact, in the continuous setting, I use the original direction in the proof.  What do we do for FSD stuff?}
\end{theorem}
\begin{proof}[Proof]
	We prove the contrapositive: if there is a convex surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F$ of codimension at most $d$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

@@

In the proof of Theorem~\ref{thm:consistent-implies-indir-elic} below, we modify the argument of~\citet[Corollary 7]{frongillo2018elicitation} to bound the elicitation complexity of the Bayes risk of the loss.
\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some $\Gamma:\simplex\to\reals^d$.
  Let $p\in\relint\simplex$ and let $\Gamma_r$ be some level set of $\Gamma$ such that (i) $p\in\Gamma_r$, (ii) $\codim(\affhull(\Gamma_r))=d+1$, and (iii) if $\Gamma_r$ is not a singleton, $\lbar$ is nonconstant on $\Gamma_r$.
  Then $\eliccvx(\lbar) \geq \min(d+1,n-1)$.
\end{theorem}
\begin{proof}
  Suppose that we have some convex loss $\hat L:\reals^k\times\Y\to\reals$ eliciting a property $\hat\Gamma:\simplex\to\reals^k$, and link $\psi : \reals^k \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  % The condition $\iden(\Gamma)=d$ implies the existence of some level set $\Gamma_r$ such that $\codim(\affhull(\Gamma_r)) \geq d$; otherwise, taking $F_r = \affhull(\Gamma_r)$ for all reports $r\in\reals^d$, we would have $\codim(F_r) \leq d-1$ and $\Gamma_r = F_r \cap \simplex$, implying $\iden(\Gamma)\leq d-1$.
  The proof of \citet[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that every level set of $\hat\Gamma$ is contained in a level set of $\Gamma$; for completeness we give the argument here.
  Suppose for a contradiction that we have $p,p'$ with $\hat\Gamma(p)=\hat\Gamma(p')$ but $\Gamma(p) \neq \Gamma(p')$.
  As $\lbar = \psi \circ \hat\Gamma$, we also have $\lbar(p) = \lbar(p')$.
  Letting $p'' = \tfrac 1 2 p +  \tfrac 1 2 p'$, Lemma~\ref{lem:elic-complex-bayes-concave} would then give us $\lbar(p'') >  \tfrac 1 2 \lbar(p) +  \tfrac 1 2 \lbar(p') = \lbar(p)$.
  By \citet{osband1985providing}, the level sets $\hat\Gamma_{\hat r}$ are convex, giving $\hat\Gamma(p'') = \hat\Gamma(p)$, which would imply $\lbar(p'')=\lbar(p)$, contradicting $\lbar = \psi \circ \hat\Gamma$.
  We conclude $\hat\Gamma$ must refine $\Gamma$, and thus $\hat L$ actually indirectly elicits $\Gamma$ through some other link function.
  % The proof of \cite[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that for all $\hat r \in \hat\Gamma(\simplex)$ we have $\hat\Gamma_{\hat r} \subseteq \Gamma_r$ for some $r\in\reals^k$.

  By Theorem~\ref{thm:cvx-flats}, as $\hat L$ elicits $\hat\Gamma$ and indirectly elicits $\Gamma$, there is a level set $\hat\Gamma_{\hat r}$ and flat $\hat F$ containing $p$ such that $\codim(\hat F) \leq k$ and $S := \hat F \cap \simplex \subseteq \hat\Gamma_{\hat r} \subseteq \Gamma_r$.
  \raft{This last statement does follow from the proof of our main theorem, but it takes a second to realize that we show the chain of inclusions.}
  Let $F = \affhull(\Gamma_r)$ and recall $\codim(F)=d+1$.
  Let $F' = \hat F \cap \affhull(\simplex)$, and observe the following: $p\in F'\cap\relint\simplex$, $F' \subseteq \affhull(\simplex)$, $S=F'\cap\simplex\subseteq\Gamma_r$, and $\codim(F') \leq k+1$.
  From the first two observations, Lemma~\ref{lem:affhull-relint} gives $\affhull(S) = \affhull(F'\cap \simplex) = F'$.
  Now as $S \subseteq \Gamma_r$, we also have $F' = \affhull(S) \subseteq \affhull(\Gamma_r) = F$,
  implying $\codim(F') \geq \codim(F)$.

  If $\Gamma_r = \{p\}$ is a singleton, then $F = \affhull(\Gamma_r) = \{p\}$, and in particular $n = \codim(F) \leq \codim(F')$, so we must have $k=d=n-1$.
  If $\Gamma_r$ is not a singleton, suppose for a contradiction that $k\leq d$.
  Then $\codim(F) = d+1 \geq k+1 \geq \codim(F')$, so we must have $\codim(F')=d+1=\codim(F)$; combining with $F' \subseteq F$ gives $F' = F$.
  We now have $S = F' \cap \simplex = F \cap \simplex = \Gamma_r$, and as $S \subseteq \hat\Gamma_{\hat r} \subseteq \Gamma_r$, we conclude $S = \hat\Gamma_{\hat r} = \Gamma_r$.
  By assumption, $\lbar$ is non-constant on $\Gamma_r$, so we have distributions $p,p' \in \Gamma_r = \hat\Gamma_{\hat r}$ with $\lbar(p)\neq\lbar(p')$, which contradicts $\lbar = \psi \circ \hat\Gamma$.
\end{proof}
